---
title: "Hierarchical Mesoscale"
author: "Erin Silver"
date: "2026-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(network)
library(igraph)
library(abind)
library(CVXR)
```

```{r}
# Write a network generator for the Stochastic Blockmodel

# Take rxr matrix of communities' edge probabilities and labeled n-vector membership list
# Generate adjacency matrix

sbm <- function(edge_probs, NODE_MEMBERS) {
  adj_matrix <- matrix(0, nrow = length(NODE_MEMBERS), ncol = length(NODE_MEMBERS))
  
  for (i in 1:length(NODE_MEMBERS)) {
    if (i+1 <= length(NODE_MEMBERS)){
      for (j in (i+1):length(NODE_MEMBERS)) {
        adj_matrix[i,j] = rnorm(1, 1, mean = edge_probs[NODE_MEMBERS[i],NODE_MEMBERS[j]])
      }
    }
  }
  
  adj_matrix[lower.tri(adj_matrix)] = t(adj_matrix)[lower.tri(adj_matrix)]
  return(adj_matrix)
}
```

```{r}
# Given a set of multiresolution communities, create some genuine differences in the structure.

# Take one n-vector membership list. Take also a 3xr assignment matrix that relates the bottom level communities to the top level communities at each level.
# This will be used to devise the edge probabilities.

assign_matrix <- function(bl_communities) {
  ml_communities = ceiling(bl_communities / 2)
  tl_communities = ceiling(ml_communities / 2)
  
  return(rbind(tl_communities, ml_communities, bl_communities))
}
```

```{r}
# Create edge probabilities. 

bl_communities = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)
assign_mrx = assign_matrix(bl_communities)

edge_probs = matrix(0, nrow = length(bl_communities), ncol = length(bl_communities))

for (i in 1:length(unique(assign_mrx[1,]))) {
  for (j in i:length(unique(assign_mrx[1,]))) {
    edge_probs[(4*i - 3):(4*i),(4*j - 3):(4*j)] = (sample(10,1) - 1)/10
  }
}

for (i in 1:length(unique(assign_mrx[2,]))) {
  for (j in i:length(unique(assign_mrx[2,]))) {
    edge_probs[(2*i - 1):(2*i),(2*j - 1):(2*j)] = edge_probs[(2*i - 1):(2*i),(2*j - 1):(2*j)] + (sample(10,1) - 1)/100
  }
}

for (i in 1:length(unique(assign_mrx[3,]))) {
  for (j in i:length(unique(assign_mrx[3,]))) {
    edge_probs[i,j] = edge_probs[i,j] + (sample(10,1) - 1)/1000
  }
}

edge_probs[lower.tri(edge_probs)] = t(edge_probs)[lower.tri(edge_probs)]
edge_probs = log(edge_probs) - log(1 - edge_probs)
```

```{r}
# Create some meaningful differences in alt edge probs
alt_edge_probs = edge_probs
alt_edge_probs[5:8,5:8] = sample(1000,1)/1000
alt_edge_probs[11:12, 9:10] = alt_edge_probs[9:10, 11:12] = sample(1000,1)/1000
```

```{r}
m = 100
NODE_MEMBERS = sample(16, 100, replace = T)
SBM_BASE = NULL
SBM_ALT = NULL 
for (i in 1:m) {
  SBM_BASE = abind(SBM_BASE, sbm(edge_probs, NODE_MEMBERS), along = 3)
  SBM_ALT = abind(SBM_ALT, sbm(alt_edge_probs, NODE_MEMBERS), along = 3)
}
```

```{r}
# For S a 2-column matrix of indices, m the number of networks per sample, A1, A2 an array of matrices

e_value <- function(S, A1, A2) {
  m <- dim(A1)[3]
  A1_bar <- apply(A1, c(1,2), mean)
  A2_bar <- apply(A2, c(1,2), mean)
  return(m / (2 * dim(S)[1]) * sum((A1_bar[S] - A2_bar[S])^2))
}
```

```{r}
# For l a level 1:3 where 3 are the top-level, a, b node groups, and given the node membership vector, return S a 2-column matrix of indices

edge_subset <- function(l, a, b, NODE_MEMBERS) {
  # Map to different resolutions
  a = (2^l*a - (2^l - 1)):2^l*a
  b = (2^l*b - (2^l - 1)):2^l*b
  nodes_a = which(NODE_MEMBERS %in% a)
  nodes_b = which(NODE_MEMBERS %in% b)
  return(as.matrix(expand.grid(nodes_a, nodes_b)))
}

# Wrapper to return e value of A in A.set

e.A <- function(A) {
  e_vals <- NULL
  for (j in A[3:6]) {
    e_vals <- append(e_vals, e_value(edge_subset(A[,2], J_SET[j,1], J_SET[j,2], NODE_MEMBERS), SBM_BASE, SBM_ALT))
  }
  mean(e_vals)
}
```

```{r}
e_value(edge_subset(1, 2, 2, NODE_MEMBERS), SBM_BASE, SBM_ALT)
```

### eLP

```{r}
library(lpSolve)
```


Per *Gablenz \& Sabatti*, we seek a vector $\mathbf x_A$ which maximizes $\sum_{A \in \mathcal A} w(A)\mathbf x_A$ subject to $x_A \in \{0,1\}$ and $|\mathcal A| - \alpha e_A\sum_Ax_A \leq |\mathcal A| \times (1 - x_A)$ and $\sum_{A;j}x_A\leq 1$. 

Where $|\mathcal A|$ is the number of hypotheses. Those will be the unique 3-tuples $(l, b, a)$. In this case, that's $\binom 4 2 + \binom 8 2 + \binom {16} 2 + 4 + 8 + 16 = 182$.

```{r}
# alpha
alpha <- 0.05

# l, a, b (full)
# J_SET <- expand.grid(1, 1:16, 1:16)
# J_SET <- J_SET[J_SET[,2] <= J_SET[,3],]

J_SET <- cbind(1:16)

# |J_SET|
length(J_SET)
```

```{r}
# For j in J_SET
i = 1
j <- J_SET[i]
```

```{r}
# Now devise A.sets:
A.set <- rbind(expand.grid(1:3, 1:4), expand.grid(1:2, 5:8), expand.grid(1, 9:16))
A.set <- cbind(A.set, matrix(0, nrow = dim(A.set)[1], ncol = 4))

for (i in 1:dim(A.set)[1]) {
    A.set[i,3:6] <- matrix((A.set[i,2]) * 2^(A.set[i,1] - 1) - 3:0 %/% 2^(3 - A.set[i,1]))
}
```

```{r}
#e.A 
e.A(A)
```
Recall:
$x_A \in \{0,1\}$ and $|\mathcal A| - \alpha e_A\sum_Ax_A \leq |\mathcal A| \times (1 - x_A)$ and $\sum_{A;j}x_A\leq 1$.

```{r}
# Suppose x.A is a vector of length |A|
# Devise constraints:

const.7b <- function(x, A) {
  x[A.set == A] == 0 | x[A.set == A] == 1
}

const.7c <- function(x, A) {
  dim(A.set)[1] - alpha * e.a(A) * sum(x) <- dim(A.set)[1] * (1 - x[A.set == A])
}

const.7d <- function(x) {
  for (j in J_SET) {
    if (sum(x[A.set[,3] == j | A.set[,4] == j | A.set[,5] == j | A.set[,6] == j] > 1)) {
      return(FALSE)
    }
  }
  TRUE
}


```

